---
title: "Data 607 Assignment 7"
author: "Jayden Jiang"
date: "2024-12-10"
output: html_document
---

```{r}
library(rjson)
library(RCurl)
library(rvest)
library(selectr)
library(httr)
library(XML)
library(stringr)
library(arrow)
library(ROAuth)
library(gt)
```

## Introduction 
This is a warm up exercise to get familiar with the JSON, HTML, XML, and Parquet file formats, and using packages to read these data formates for downstream use in R data frames. Use data below for inventory analysis at the retailer. Preparing the data for analysis with listed formats, provide the pros and cons of each formats. 

## JSON Format
```{r}
# Create CUNYMart data
cunymart_data_json <- list(
  list(Category = "Electronics", Item_Name = "Smartphone", Item_ID = 101, Brand = "TechBrand", Price = 699.99,
       Variation_ID = "101-A", Variation_Details = "Color: Black, Storage: 64GB"),
  list(Category = "Electronics", Item_Name = "Smartphone", Item_ID = 101, Brand = "TechBrand", Price = 699.99,
       Variation_ID = "101-B", Variation_Details = "Color: White, Storage: 128GB"),
  list(Category = "Electronics", Item_Name = "Laptop", Item_ID = 102, Brand = "CompuBrand", Price = 1099.99, 
       Variation_ID = "102-A", Variation_Details = "Color: Silver, Storage: 256GB")
)

# Export JSON data
json_format <- toJSON(cunymart_data_json)
write(json_format, file = "cunymart_data.json")

# Import JSON file
imported_json <- fromJSON(file = "cunymart_data.json")
imported_json
```

## HTML Format
```{r}
# Create CUNYMart data as data frame
cunymart_data_html <- data.frame(
  Category = c("Electronics", "Electronics", "Electronics"),
  Item_Name = c("Smartphone", "Smartphone", "Laptop"),
  Item_ID = c(101, 101, 102),
  Brand = c("TechBrand", "TechBrand", "CompuBrand"),
  Price = c(699.99, 699.99, 1099.99),
  Variation_ID = c("101-A", "101-B", "102-A"),
  Variation_Details = c("Color: Black, Storage: 64GB",
                        "Color: White, Storage: 128GB",
                        "Color: Silver, Storage: 256GB")
)

# Export as HTML using HTML table
html_table <- paste0('<table border="1">',
                     paste0('<tr>', paste0('<th>', names(cunymart_data_html), '</th>'), '</tr>'),
                     paste0(
                       apply(cunymart_data_html, 1, function(row) paste0('<tr>', paste0('<td>', row, '</td>'), '</tr>')),
                       collapse = "\n"
                     ), '</table>')
write(html_table, file = "cunymart_data.html")

# Import HTML file using rvest
html_content <- read_html("cunymart_data.html")
extracted_table <- html_table(html_content)
extracted_table
```

## XML Format
```{r}
# Create XML document from scratch
xml_doc <- newXMLDoc()
root_node <- newXMLNode("Inventory", doc = xml_doc)

# Add items to XML
items <- list(
  list(Category = "Electronics", Item_Name = "Smartphone", Item_ID = 101, Brand = "TechBrand", Price = 699.99,
       Variation_ID = "101-A", Variation_Details = "Color: Black, Storage: 64GB"),
  list(Category = "Electronics", Item_Name = "Smartphone", Item_ID = 101, Brand = "TechBrand", Price = 699.99,
       Variation_ID = "101-B", Variation_Details = "Color: White, Storage: 128GB")
)

for (item in items) {
  item_node <- newXMLNode("Item", parent = root_node)
  for (key in names(item)) {
    newXMLNode(key, item[[key]], parent = item_node)
  }
}

# Export to XML file
saveXML(xml_doc, file = "cunymart_data.xml")

# Import XML back into R
imported_xml <- xmlParse("cunymart_data.xml")
xmlToList(imported_xml)
```

## Parquet Format 
```{r}
# Create CUNYMart data
cunymart_data_parquet <- data.frame(
  Category = c("Electronics", "Electronics", "Electronics"),
  Item_Name = c("Smartphone", "Smartphone", "Laptop"),
  Item_ID = c(101, 101, 102),
  Brand = c("TechBrand", "TechBrand", "CompuBrand"),
  Price = c(699.99, 699.99, 1099.99),
  Variation_ID = c("101-A", "101-B", "102-A"),
  Variation_Details = c("Color: Black, Storage: 64GB",
                        "Color: White, Storage: 128GB",
                        "Color: Silver, Storage: 256GB")
)

# Export to Parquet
write_parquet(cunymart_data_parquet, "cunymart_data.parquet")

imported_parquet <- read_parquet("cunymart_data.parquet")
head(imported_parquet)
```

## Summary Table 
```{r}
pros_cons <- data.frame(
  Format = c("JSON", "HTML", "XML", "Parquet"),
  Pros = c("Simple to debug, Easy to read & write, Used in APIs",
           "Great for displaying in web browsers, Interative visualizations",
           "Hierarchial data support, self-descriptive format",
           "Highly compressed, Fast for big data and analytics"),
  Cons = c("No Metadata, Harder to work with large files",
           "Not a suitable data exchange format, Parsing is harder",
           "Verbose and large file sizes, Harder to read",
           "Harder to read, Complex to debug")
)

proscons_table <- gt(pros_cons) %>%
  tab_header(
    title = "Pros and Cons of Different Data Formats"
  )

proscons_table
```



